---
title: "intro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates how to use this package to benchmark NMF runtimes.

```{r setup}
library(FastNMF)
```

We will use `microbenchmark` for benchmarking, `RcppML` for an example dataset, and compare runtime of the `FastNMF::FastNMF` function vs. the RcppML `nmf` function.

You will need to use `install.packages(RcppML)`, for example, to install these libraries on your system:

```{R}
library(RcppML)
library(microbenchmark)
library(ggplot2)
data(hawaiibirds) # provided in the RcppML package
```

Load the example `hawaiibirds` dataset and learn an NMF model at a rank of 10:

```{R}
set.seed(123)
nmf_model <- FastNMF(hawaiibirds$counts, k = 10)
mb <- microbenchmark(
  FastNMF(hawaiibirds$counts, k = 10), 
  RcppML::nmf(hawaiibirds$counts, k = 10), 
  times = 10)
mb
```

And plot the benchmarking results:

```{R, fig.width = 6, fig.height = 3}
autoplot(mb)
```

Note that runtime is incredibly dependent on dataset size, shape, sparsity, and factor rank.  Benchmarking should demonstrate clear differences in runtime across a wide range of dataset characteristics.

Also be aware of local compiler optimizations and local Makevars files vs. optimizations used by central repositories or `devtools::install_github("FastNMF")`.
